### 2018-12-10 Manuscript/sup_mat/sup_mat.Rmd
Updated Sup_mat
Added a table and new numbers for the trials etc with the most up to date data

---
title: "Supplementary Materials"
output:
  bookdown::html_document2:
    toc: yes
    toc_depth: 2
    fig_caption: yes
---

```{r setup, include=FALSE}
# setting options
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(knitr)

# what data do we need?
# get line seg acc and rt 
ls_acc_RT <- readRDS(file = "../../Scripts/lineseg/scratch/processedRTandAccData.Rda")

```

# Participants
```{r Participants, echo = FALSE}
# sort out participants 
ls_acc_RT$observer <- as.numeric(as.factor(as.numeric(ls_acc_RT$observer)))

# number of participants used
num_participants <- length(unique(ls_acc_RT$observer))

```

```{r processing ls data, echo = FALSE, include = FALSE}
# load in unprocessed data 
load("../../Scripts/lineseg/scratch/unfiltered_fixDat")
load("../../Scripts/lineseg/scratch/unfiltered_trlDat")
trlDat_unf <- trlDat
# tidy 
rm(trlDat)

# sort out participant levels 
fixDat$observer <- as.numeric(as.factor(as.numeric(fixDat$observer)))

fixDat_unf <- fixDat
# tidy 
rm(fixDat)

#### getting accuracy and mistaken keys ####
invalid_kp <- trlDat_unf %>%
  group_by(responseKey) %>%
  summarise(n = n()) %>%
  filter(responseKey == "Key_x")
invalid_kp <- invalid_kp$n

# filter out bad participants 
avg_acc_all <- ls_acc_RT %>%
  filter(observer != 15,
         observer != 25,
         observer != 47,
         observer != 48) %>%
  group_by(observer, targSide) %>%
  summarise(Acc = mean(accuracy))

# min acc easy 
min_easy <- round(min(avg_acc_all$Acc[avg_acc_all$targSide == "easy"]), digits = 3)*100

# min acc absent 
min_absent <- round(min(avg_acc_all$Acc[avg_acc_all$targSide == "absent"]), digits = 3)*100

#### sorting out fixation data #### 
num_fixations <- nrow(fixDat_unf)

# get removed fixations data 
load("../../Scripts/lineseg/scratch/processedFixationData")
# get difference value 
num_removed <- num_fixations - nrow(fixDat)

```
Data was collected from 64 Participants as originally planned. Four of these participants did not complete all parts of the experiment (either declining to participate in the second session, or could not be calibrated with the eye tracker), so four new participants were recruited to bring the total back up to `r num_participants`.

## Split-Half
Accuracy data is shown in Figure \ref{fig:fig1}. From this we can see that there are a number of outliers: 15, 25, 47, and 48 (NB: these may have changed with new labelling). These participants, in at least one session, either missed the majority of easy targets, or responded with false positives on the majority of target absent trials. After removing these participants, the lowest accuracy in either session was `r paste(min_easy,"%", sep = "")` for the easy targets, and `r paste(min_absent, "%", sep = "")` for target absent trials. This leaves us with `r length(unique(avg_acc_all$observer))` participants for the split-half paradigm.

```{r fig1, fig.cap = "\\label{fig1} Accuracy data for each participant for the split-half paradigm. ", echo = FALSE, fig.align = "center"}
# make the plot
plt <- ls_acc_RT %>%
  group_by(observer, targSide, session) %>%
  summarise(Accuracy = mean(accuracy)) %>%
  ggplot(aes(targSide, Accuracy, fill = session)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_y_continuous(breaks = c(0,1)) + 
  theme_bw() +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5), legend.position = "top", strip.text.x = element_text(margin = margin(0.01,0,0.01,0, "mm"))) + 
  scale_fill_ptol() +
  facet_wrap(~observer)
plt$labels$x <- "Target Condition"
plt
```

```{r fig2, fig.cap = "\\label{fig2} Distribution of reaction times in split-half paradigm.", echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
plt <- ls_acc_RT %>%
  ggplot(aes(rt)) + 
  geom_histogram() + 
  theme_bw() +  
  facet_wrap(~targSide)
plt$labels$x <- "Response Time (msecs)"
plt2 <- ls_acc_RT %>%
  mutate(log2_RT = log2(rt)) %>%
  ggplot(aes(log2_RT)) + 
  geom_histogram() + 
  theme_bw() +
  facet_wrap(~ targSide)
plt2$labels$x <- "log2 Response Time"

# put together 
grid.arrange(plt,plt2)
```

# Data Processing

## Split-Half 
`r invalid_kp` trials with invalid key responses were removed. After removing data from the four outlier participants (see above), all remaining incorrect trials `r #find out what this number is meant to be` (n = 2529) were removed, leaving a total of `r nrow(trlDat_unf) - invalid_kp` trials over 60 participants.

### Reaction Times 
As expected, reaction times were highly skewed (Figure 2), so were log2 transformed (in ms units) before the participant means were calculated for each session and target condition (Figure )

### Eye Movements
`r num_fixations` fixations were recorded. Of these, `r num_removed` fell outside of the stimuli area and were removed. Fixations landing within a vertical strip consisting of 10% of the stimuli's width were classed as central. All remaining fixations were then classed as landing on the homogeneous or heterogeneous half of the stimulus. Initial fixations were not included in the analysis. Numbers of fixations are given in Table 2.

```{r make table, echo = FALSE, results='asis'}
# get fixation location
fixDat <- fixDat %>%
  mutate(fix_loc = "central")
fixDat$fix_loc[fixDat$x < -0.1 & fixDat$easySide == "left"] <- "homogenous"
fixDat$fix_loc[fixDat$x > 0.1 & fixDat$easySide == "right"] <- "homogenous"
fixDat$fix_loc[fixDat$x < -0.1 & fixDat$easySide == "right"] <- "heterogenous"
fixDat$fix_loc[fixDat$x > 0.1 & fixDat$easySide == "left"] <- "heterogenous"

# get num fixations in these locations 
table_dat_all <- fixDat %>%
  group_by(targSide, fix_loc) %>%
  filter(n < 6) %>%
  summarise(number = n()) %>%
  mutate(range = "1 < n")

table_dat_2_to_5 <- fixDat %>%
  group_by(targSide, fix_loc) %>%
  filter(n >= 2,
         n <= 5) %>%
  summarise(number = n()) %>%
  mutate(range = "2 <= n <= 5")

table_dat <- rbind(table_dat_all, table_dat_2_to_5) %>%
  spread(range, number)

# make table 
kable(table_dat, caption = "table thing")
```

## Adaptive Choice

## Foraging